---
title: "NHANES reproduction bioage collaboration"
author: "Talia + Calen + Waylon (+ others?)"
date: "November 11, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load packages
```{r message=FALSE, warning=FALSE }
library(RNHANES)
library(haven)
library(psych)
library(dplyr)
library(ggplot2)
library(survey)
library(factoextra)
library(jtools)
library(ggstance)
library(corrplot)
library(GGally)
library(stargazer)
```

##Download data, merge stuff, & do some transformations
```{r message=FALSE}
demo1 <- nhanes_load_data("DEMO", "1999-2000", demographics = FALSE)
demo2 <- nhanes_load_data("DEMO_B", "2001-2002", demographics = FALSE)

demo1 <- demo1[,c("SEQN", "SDMVPSU", "SDMVSTRA")]
demo2 <- demo2[,c("SEQN", "SDMVPSU", "SDMVSTRA")]
DEMO <- rbind(demo1,demo2)

#### Import smoking aq's

smq1 <- nhanes_load_data("SMQ", "1999-2000", demographics = FALSE)
smq2 <- nhanes_load_data("SMQ_B", "2001-2002", demographics = FALSE)

smq1_short<-smq1[c("SMQ020", "SMQ040", "SEQN")]
smq2_short<-smq2[c("SMQ020", "SMQ040", "SEQN")]
#SMQ020 Smoked 100 cigarettes in life? 1 = yes 2 = no
#SMQ040 DO you now smoke cigarettes? 1 = every day, 2 = some days, 3 = not at all

SMQ<- rbind(smq1_short,smq2_short)

SMQ <- na_if(SMQ, 7)
SMQ <- na_if(SMQ, 9)

SMQ$smoking <- 2 #current smoker
SMQ$smoking[SMQ$SMQ020==2]<-0 #never smoker
SMQ$smoking[SMQ$SMQ040==3]<-1 #past smoker

#remove one person with no SEQN
SMQ <- SMQ[which(SMQ$SEQN>0),]

#### RHQ 

rhq1 <- nhanes_load_data("RHQ", "1999-2000", demographics = FALSE)
rhq2 <- nhanes_load_data("RHQ_B", "2001-2002", demographics = FALSE)

rhq1_short<- rhq1[c("SEQN", "RHD130", "RHQ140", "RHQ150", "RHQ160", "RHD170", "RHQ180", "RHQ190", "RHQ200", "RHQ210", "RHD230","RHQ030", "RHQ040", "RHQ060")]
#RHD130 Ever been pregnant? 1 = yes, 2= no
#RHQ140 Think you are pregnant now? 1 = yes, 9 = don't know
#RHQ150 What month of pregnancy are you in? 1-10 = range of values, 99 = don't know
#RHQ160 How many times have you been pregnant? 1-55 = range of values, 77 = refused, 99 = don't know
#RHD170 Number of pregnancies resulting in live births 1-10 = range of values, 11 = 11 or more, 77 = refused
#RHQ180 Age of first live birth 13-41 = range of values, 999= don't know
#RHQ190 Age at last live birth 11-57 = range of values, 999 = don't know
#RHQ200 Now breastfeeding a child? 1 = yes, 2 = no
#RHQ210 breastfed any of your children? 1 = yes, 2 = no
#RHD230 Number of children breastfed at least 1 month 0-13 = range of values
#RHQ030 had regular periods in the last 12 months?
#RHQ040 reason not having regular periods (5= menopause)

rhq2_short<- rhq2[c("SEQN", "RHD130", "RHQ141", "RHQ151", "RHQ160", "RHD170", "RHQ180", "RHQ190", "RHQ200", "RHQ210", "RHD230","RHQ030", "RHQ040", "RHQ060")]
#RHD130 Ever been pregnant? 1 = yes, 2= no, 9 = don't know
#RHQ141 Think you are pregnant now? 1 = yes, 9 = don't know
#RHQ151 What month of pregnancy are you in? 1-9 = range of values, 99 = don't know
#RHQ160 How many times have you been pregnant? 1-18 = range of values, 77 = refused, 99 = don't know
#RHD170 Number of pregnancies resulting in live births 1-10 = range of values, 11 = 11 or more, 77 = refused
#RHQ180 Age of first live birth 12-39 = range of values, 999= don't know
#RHQ190 Age at last live birth 13-57 = range of values, 999 = don't know
#RHQ200 Now breastfeeding a child? 1 = yes, 2 = no
#RHQ210 breastfed any of your children? 1 = yes, 2 = no, don't know = 9
#RHD230 Number of children breastfed at least 1 month 0-13 = range of values, 99 = don't know

#Change column names so can rbind
colnames(rhq2_short)[3:4] <- c("RHQ140","RHQ150")

#Combine cycles
RHQ<- rbind(rhq1_short,rhq2_short)

#Change column names so they don't suck
colnames(RHQ)[2:14] <- c("everpreg", "currentpreg", "currentpregmonth", "timespreg", "livebirths", "firstlivebirth", "lastlivebirth", "currentbreastfeed", "everbreastfed", "numberbreastfed", "mensregularity", "irregreason", "ageLMP")

#Recode the 'refused' and 'don't know's in RHQ data as NA (values differ by variable)
RHQ <- na_if(RHQ, 999)
RHQ <- na_if(RHQ, 77)
RHQ <- na_if(RHQ, 99)
RHQ$everpreg <- na_if(RHQ$everpreg, 9)
RHQ$everbreastfed <- na_if(RHQ$everbreastfed, 9)

RHQSMQ <- merge(RHQ, SMQ, by = "SEQN", all = TRUE)
NHANES <- merge(RHQSMQ, DEMO, by = "SEQN", all = TRUE)

#Create quadratic term for live births
NHANES$livebirths2 <- NHANES$livebirths^2

#Transform the variable indexing whether a woman has ever been pregnant so that 0 = never pregnant, and 1 = ever pregnant
NHANES$everpreg[NHANES$everpreg==2]<-0

#Download biological age file
NHANES_BA_Talia <- read_sav(here("Data/NHANES_BA_Talia.sav"))
colnames(NHANES_BA_Talia)[1] <- "SEQN"

#Combine all data
df <- merge(NHANES,NHANES_BA_Talia,by="SEQN")

#Recode race/ethnicity to group all Hispanic-identifying individuals together
df$ridreth1[df$ridreth1==3]<-1

#Create variable for menopause
df$menopause <- NA
df$menopause[df$irregreason==5] <- 1
df$menopause[df$irregreason==1 | df$irregreason==2 | df$irregreason==3 | df$irregreason==4 | df$irregreason==6] <- 0
df$timenoperiod<-df$age-df$ageLMP
df$menopause[df$timenoperiod >1 & (df$irregreason!=1 | df$irregreason!=2 | df$irregreason!=3 | df$irregreason!=4 | df$irregreason!=6 | df$irregreason==5)] <- 1
df$menopause[df$age>61] <- 1
df$menopause[df$age<41] <- 0
df$menopause[df$Gender==2]<-NA
df$menopause[df$mensregularity==1] <- 0
```

##Specify survey designs: # of live births
```{r message=FALSE}
base <- svydesign(id      = ~SDMVPSU,
                          strata  = ~SDMVSTRA,
                          weights = ~WTMEC4YR,
                          nest    = TRUE,
                          data    = df)

nhanesDesign1<- subset(base, age > 17 & pregnant != 1 & livebirths > -1 & menopause == 0)
nhanesDesign2<- subset(base, age > 17 & pregnant != 1 & livebirths > -1 & menopause == 1)
```

##Most important model (all covariates, no interactions): Premenopausal, # live births
```{r message=FALSE, echo=T, results='hide'}
summary(a1 <- svyglm(HD ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
summary(a2 <- svyglm(LM ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
summary(a3 <- svyglm(KDM ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))      
summary(a4 <- svyglm(log_telo ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
```

##Tables for premenopausal women, # live births
```{r, results = 'asis', echo = FALSE}
stargazer(a1,a2,a3,a4, single.row = F, align = T, intercept.bottom = T, no.space = T, dep.var.caption = "", type = "html", out = "models.html", report = "vc*",   star.char = c("+", "*", "**", "***"), digits = 3, notes = "+p<0.1;*p<0.05;**p<0.01;***p<0.001 ", notes.append = FALSE)
```

##Most important model (all covariates, no interactions): Postmenopausal, # live births
```{r message=FALSE, echo=T, results='hide'}
summary(b1 <- svyglm(HD ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
summary(b2 <- svyglm(LM ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
summary(b3 <- svyglm(KDM ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))      
summary(b4 <- svyglm(log_telo ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
```

##Tables for postmenopausal women, # live births
```{r, results = 'asis', echo = FALSE}
stargazer(b1,b2,b3,b4, single.row = F, align = T, intercept.bottom = T, no.space = T, dep.var.caption = "", type = "html", out = "models.html", report = "vc*",   star.char = c("+", "*", "**", "***"), digits = 3, notes = "+p<0.1;*p<0.05;**p<0.01;***p<0.001 ", notes.append = FALSE)
```

##Plots, # live births (model 1 = premenopausal, model 2 = postmenopausal)
```{r message=FALSE, echo=FALSE}

apatheme=theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        axis.line=element_line(),
        legend.title=element_blank(), 
        axis.text=element_text(size=20),
        axis.title=element_text(size=20),
        legend.text = element_text(size = 16),
        axis.title.y = element_blank(),
        plot.title = element_text(size = 20, face = "bold"))

#Plot for HD
HDplot1<-plot_summs(a1, b1,  
           coefs = c("# Live Births (linear)" = "livebirths", "# Live births (quadratic)" = "livebirths2", "Current Age" = "age"),
           scale = TRUE) + apatheme + ggtitle("Homeostatic Dysregulation")

#Plot for LM
LMplot1<-plot_summs(a2, b2,
                   coefs = c("# Live Births (linear)" = "livebirths", "# Live births (quadratic)" = "livebirths2", "Current Age" = "age"),
                   scale = TRUE) + apatheme + ggtitle("Levine Method")

#Plot for KDM
KDMplot1<-plot_summs(a3, b3,
                   coefs = c("# Live Births (linear)" = "livebirths", "# Live births (quadratic)" = "livebirths2", "Current Age" = "age"),
                   scale = TRUE) + apatheme + ggtitle("Klemera-Doubal Method")

#Plot for telo
TELOplot1<-plot_summs(a4, b4, 
                    coefs = c("# Live Births (linear)" = "livebirths", "# Live births (quadratic)" = "livebirths2", "Current Age" = "age"),
                    scale = TRUE) + apatheme + ggtitle("Telomere Length (log)")

HDplot1
LMplot1
KDMplot1
TELOplot1
```

##Specify survey designs: ever pregnant
```{r message=FALSE}
base <- svydesign(id      = ~SDMVPSU,
                          strata  = ~SDMVSTRA,
                          weights = ~WTMEC4YR,
                          nest    = TRUE,
                          data    = df)

nhanesDesign1<- subset(base, age > 17 & pregnant != 1 & everpreg > -1 & menopause == 0)
nhanesDesign2<- subset(base, age > 17 & pregnant != 1 & everpreg > -1 & menopause == 1)
```

##Most important model (all covariates, no interactions): Premenopausal, ever pregnant
```{r message=FALSE, echo=T, results='hide'}
summary(c1 <- svyglm(HD ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
summary(c2 <- svyglm(LM ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
summary(c3 <- svyglm(KDM ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))      
summary(c4 <- svyglm(log_telo ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
```

##Tables for premenopausal women, ever pregnant
```{r, results = 'asis', echo = FALSE}
stargazer(c1,c2,c3,c4, single.row = F, align = T, intercept.bottom = T, no.space = T, dep.var.caption = "", type = "html", out = "models.html", report = "vc*",   star.char = c("+", "*", "**", "***"), digits = 3, notes = "+p<0.1;*p<0.05;**p<0.01;***p<0.001 ", notes.append = FALSE)
```

##Most important model (all covariates, no interactions): Postmenopausal, ever pregnant
```{r message=FALSE, echo=T, results='hide'}
summary(d1 <- svyglm(HD ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
summary(d2 <- svyglm(LM ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
summary(d3 <- svyglm(KDM ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))      
summary(d4 <- svyglm(log_telo ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
```

##Tables for postmenopausal women, ever pregnant
```{r, results = 'asis', echo = FALSE}
stargazer(d1,d2,d3,d4, single.row = F, align = T, intercept.bottom = T, no.space = T, dep.var.caption = "", type = "html", out = "models.html", report = "vc*",   star.char = c("+", "*", "**", "***"), digits = 3, notes = "+p<0.1;*p<0.05;**p<0.01;***p<0.001 ", notes.append = FALSE)
```

##Plots, ever pregnant (model 1 = premenopausal, model 2 = postmenopausal)
```{r message=FALSE, echo=FALSE}
#Plot for HD
HDplot2<-plot_summs(c1, d1, 
                   coefs = c("Parity" = "everpreg", "Current Age" = "age"), 
                   scale = TRUE) + apatheme + ggtitle("Homeostatic Dysregulation")

#Plot for LM
LMplot2<-plot_summs(c2, d2,
                   coefs = c("Parity" = "everpreg", "Current Age" = "age"),
                   scale = TRUE) + apatheme + ggtitle("Levine Method")

#Plot for KDM
KDMplot2<-plot_summs(c3, d3,  
                    coefs = c("Parity" = "everpreg", "Current Age" = "age"),
                    scale = TRUE) + apatheme + ggtitle("Klemera-Doubal Method")

#Plot for telo
TELOplot2<-plot_summs(c4, d4, 
                     coefs = c("Parity" = "everpreg", "Current Age" = "age"), 
                     scale = TRUE) + apatheme + ggtitle("Telomere Length (log)")

HDplot2
LMplot2
KDMplot2
TELOplot2
```