---
title: "NHANES reproduction bioage collaboration"
author: "Talia + Calen + Waylon (+ others?)"
date: "November 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load packages
```{r message=FALSE, warning=FALSE }
library(RNHANES)
library(haven)
library(psych)
library(dplyr)
library(ggplot2)
library(survey)
library(factoextra)
library(jtools)
library(ggstance)
library(corrplot)
library(GGally)
library(stargazer)

# install.packages(here) # Run this if you don't have here package. See below for how it works. 

library(here)
library(jtools)

```

##Download data, merge stuff, & do some transformations
```{r message=FALSE}
demo1 <- nhanes_load_data("DEMO", "1999-2000", demographics = FALSE)
demo2 <- nhanes_load_data("DEMO_B", "2001-2002", demographics = FALSE)

demo1 <- demo1[,c("SEQN", "SDMVPSU", "SDMVSTRA")]
demo2 <- demo2[,c("SEQN", "SDMVPSU", "SDMVSTRA")]
DEMO <- rbind(demo1,demo2)

#### Import smoking aq's

smq1 <- nhanes_load_data("SMQ", "1999-2000", demographics = FALSE)
smq2 <- nhanes_load_data("SMQ_B", "2001-2002", demographics = FALSE)

smq1_short<-smq1[c("SMQ020", "SMQ040", "SEQN")]
smq2_short<-smq2[c("SMQ020", "SMQ040", "SEQN")]
#SMQ020 Smoked 100 cigarettes in life? 1 = yes 2 = no
#SMQ040 DO you now smoke cigarettes? 1 = every day, 2 = some days, 3 = not at all

SMQ<- rbind(smq1_short,smq2_short)

SMQ <- na_if(SMQ, 7)
SMQ <- na_if(SMQ, 9)

SMQ$smoking <- 2 #current smoker
SMQ$smoking[SMQ$SMQ020==2]<-0 #never smoker
SMQ$smoking[SMQ$SMQ040==3]<-1 #past smoker

#remove one person with no SEQN
SMQ <- SMQ[which(SMQ$SEQN>0),]

#### RHQ 

rhq1 <- nhanes_load_data("RHQ", "1999-2000", demographics = FALSE)
rhq2 <- nhanes_load_data("RHQ_B", "2001-2002", demographics = FALSE)

rhq1_short<- rhq1[c("SEQN", "RHD130", "RHQ140", "RHQ150", "RHQ160", "RHD170", "RHQ180", "RHQ190", "RHQ200", "RHQ210", "RHD230","RHQ030", "RHQ040", "RHQ060")]
#RHD130 Ever been pregnant? 1 = yes, 2= no
#RHQ140 Think you are pregnant now? 1 = yes, 9 = don't know
#RHQ150 What month of pregnancy are you in? 1-10 = range of values, 99 = don't know
#RHQ160 How many times have you been pregnant? 1-55 = range of values, 77 = refused, 99 = don't know
#RHD170 Number of pregnancies resulting in live births 1-10 = range of values, 11 = 11 or more, 77 = refused
#RHQ180 Age of first live birth 13-41 = range of values, 999= don't know
#RHQ190 Age at last live birth 11-57 = range of values, 999 = don't know
#RHQ200 Now breastfeeding a child? 1 = yes, 2 = no
#RHQ210 breastfed any of your children? 1 = yes, 2 = no
#RHD230 Number of children breastfed at least 1 month 0-13 = range of values
#RHQ030 had regular periods in the last 12 months?
#RHQ040 reason not having regular periods (5= menopause)

rhq2_short<- rhq2[c("SEQN", "RHD130", "RHQ141", "RHQ151", "RHQ160", "RHD170", "RHQ180", "RHQ190", "RHQ200", "RHQ210", "RHD230","RHQ030", "RHQ040", "RHQ060")]
#RHD130 Ever been pregnant? 1 = yes, 2= no, 9 = don't know
#RHQ141 Think you are pregnant now? 1 = yes, 9 = don't know
#RHQ151 What month of pregnancy are you in? 1-9 = range of values, 99 = don't know
#RHQ160 How many times have you been pregnant? 1-18 = range of values, 77 = refused, 99 = don't know
#RHD170 Number of pregnancies resulting in live births 1-10 = range of values, 11 = 11 or more, 77 = refused
#RHQ180 Age of first live birth 12-39 = range of values, 999= don't know
#RHQ190 Age at last live birth 13-57 = range of values, 999 = don't know
#RHQ200 Now breastfeeding a child? 1 = yes, 2 = no
#RHQ210 breastfed any of your children? 1 = yes, 2 = no, don't know = 9
#RHD230 Number of children breastfed at least 1 month 0-13 = range of values, 99 = don't know

#Change column names so can rbind
colnames(rhq2_short)[3:4] <- c("RHQ140","RHQ150")

#Combine cycles
RHQ<- rbind(rhq1_short,rhq2_short)

#Change column names so they don't suck
colnames(RHQ)[2:14] <- c("everpreg", "currentpreg", "currentpregmonth", "timespreg", "livebirths", "firstlivebirth", "lastlivebirth", "currentbreastfeed", "everbreastfed", "numberbreastfed", "mensregularity", "irregreason", "ageLMP")

#Recode the 'refused' and 'don't know's in RHQ data as NA (values differ by variable)
RHQ <- na_if(RHQ, 999)
RHQ <- na_if(RHQ, 77)
RHQ <- na_if(RHQ, 99)
RHQ$everpreg <- na_if(RHQ$everpreg, 9)
RHQ$everbreastfed <- na_if(RHQ$everbreastfed, 9)

RHQSMQ <- merge(RHQ, SMQ, by = "SEQN", all = TRUE)
NHANES <- merge(RHQSMQ, DEMO, by = "SEQN", all = TRUE)

#Create quadratic term for live births
NHANES$livebirths2 <- NHANES$livebirths^2

#Transform the variable indexing whether a woman has ever been pregnant so that 0 = never pregnant, and 1 = ever pregnant
NHANES$everpreg[NHANES$everpreg==2]<-0

#Download biological age file
NHANES_BA_Talia <- read_sav(here::here("Data/NHANES_BA_Talia.sav"))
colnames(NHANES_BA_Talia)[1] <- "SEQN"

#Combine all data
df <- merge(NHANES,NHANES_BA_Talia,by="SEQN")

#Recode race/ethnicity to group all Hispanic-identifying individuals together
df$ridreth1[df$ridreth1==3]<-1

#Create variable for menopause
df$menopause <- NA
df$menopause[df$irregreason==5] <- 1
df$menopause[df$irregreason==1 | df$irregreason==2 | df$irregreason==3 | df$irregreason==4 | df$irregreason==6] <- 0
df$timenoperiod<-df$age-df$ageLMP
df$menopause[df$timenoperiod >1 & (df$irregreason!=1 | df$irregreason!=2 | df$irregreason!=3 | df$irregreason!=4 | df$irregreason!=6 | df$irregreason==5)] <- 1
df$menopause[df$age>61] <- 1
df$menopause[df$age<41] <- 0
df$menopause[df$Gender==2]<-NA
df$menopause[df$mensregularity==1] <- 0
```

##Specify survey designs: # of live births
```{r message=FALSE}
base <- svydesign(id      = ~SDMVPSU,
                          strata  = ~SDMVSTRA,
                          weights = ~WTMEC4YR,
                          nest    = TRUE,
                          data    = df)

nhanesDesign1<- subset(base, age > 17 & age <=84 & pregnant != 1 & livebirths > -1 & menopause == 0)
nhanesDesign2<- subset(base, age > 17 & age <=84 & pregnant != 1 & livebirths > -1 & menopause == 1)
```

##Most important model (all covariates, no interactions): Premenopausal, # live births
```{r message=FALSE, echo=T, results='hide'}
summary(a1_linear <- svyglm(HD ~ livebirths + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
summary(a2_linear <- svyglm(LM ~ livebirths + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
summary(a3_linear <- svyglm(KDM ~ livebirths + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))      
summary(a4_linear <- svyglm(log_telo ~ livebirths+ age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))


summary(a1 <- svyglm(HD ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
summary(a2 <- svyglm(LM ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
summary(a3 <- svyglm(KDM ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))      
summary(a4 <- svyglm(log_telo ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))


anova(a1, a1_linear, test = "Chisq")
anova(a2, a2_linear, test = "Chisq")
anova(a3, a3_linear, test = "Chisq")
# Either the quadratic outperforms (a1 < 0.1, a2 <0.05) or is  non-significantly different. Logic for quadratic. 

```

##Tables for premenopausal women, # live births
```{r, results = 'asis', echo = FALSE}
stargazer(a1,a2,a3,a4, single.row = F, align = T, intercept.bottom = T, no.space = T, dep.var.caption = "", type = "html", out = "../Output/Tables/models_pre_meno.html", report = "vc*",   star.char = c("+", "*", "**", "***"), digits = 3, notes = "+p<0.1;*p<0.05;**p<0.01;***p<0.001 ", notes.append = FALSE)
```


##Figures of curves (GARBAGE)
```{r, results = 'asis', echo = FALSE}
# make_predictions() # can't figure it out.
# ggpredict() # can't figure it out.


# IS this it? 
df %>% 
  filter(age > 17 & age <=84 & pregnant != 1 & livebirths > -1 & menopause == 0) %>% 
ggplot(., aes(x = livebirths, y = LM)) +
  geom_jitter(alpha = 0.4, col = "navy", width = 0.2)+
  geom_smooth(method = "glm", formula =  y ~ x + I(x^2), n = 500)+
  theme_bw()



```

##Most important model (all covariates, no interactions): Postmenopausal, # live births
```{r message=FALSE, echo=T, results='hide'}
summary(b1 <- svyglm(HD ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
summary(b2 <- svyglm(LM ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
summary(b3 <- svyglm(KDM ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))      
summary(b4 <- svyglm(log_telo ~ livebirths + livebirths2 + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
```

##Tables for postmenopausal women, # live births
```{r, results = 'asis', echo = FALSE}
stargazer(b1,b2,b3, single.row = F, align = T, intercept.bottom = T, no.space = T, dep.var.caption = "", type = "html", out = "../Output/Tables/models_post_meno.html", report = "vc*",   star.char = c("+", "*", "**", "***"), digits = 3, notes = "+p<0.1;*p<0.05;**p<0.01;***p<0.001 ", notes.append = FALSE)
```




## All Biological Aging Table
```{r}
stargazer(a1,a2,a3,b1,b2,b3, 
          single.row = F, 
          align = T, 
          intercept.bottom = T, 
          no.space = T, 
          dep.var.caption = "Biological Aging (BA) Measures",
          dep.var.labels = c("", "Pre-menopausal", "", "", "Post-menopausal", ""),
          column.labels = c("HD", "LM", "KDM", "HD", "LM", "KDM"),
          covariate.labels = c("Births", "Births2", "Age",
                               "BMI", "Smoking (Y/N)", "Education",
                               "Poverty", "RIDRETH", "Food Security", "Intercept"),
          type = "html", 
          out = "../Output/Tables/models_all_meno.html", 
          report = "vc*",   
          star.char = c("+", "*", "**", "***"), 
          digits = 3, 
          notes = "+p<0.1;*p<0.05;**p<0.01;***p<0.001 ", 
          notes.append = FALSE)

```




##Plots, # live births (model 1 = premenopausal, model 2 = postmenopausal)
```{r message=FALSE, echo=FALSE}

apatheme=theme_bw()+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        panel.border=element_blank(),
        axis.line=element_line(),
        legend.title=element_blank(), 
        axis.text=element_text(size=20),
        axis.title=element_text(size=20),
        legend.text = element_text(size = 16),
        axis.title.y = element_blank(),
        plot.title = element_text(size = 20, face = "bold"))

#Plot for HD
HDplot1<-plot_summs(a1, b1,  
           coefs = c("# Live Births (linear)" = "livebirths", "# Live births (quadratic)" = "livebirths2", "Current Age" = "age"),
           scale = TRUE) + apatheme + ggtitle("Homeostatic Dysregulation")

#Plot for LM
LMplot1<-plot_summs(a2, b2,
                   coefs = c("# Live Births (linear)" = "livebirths", "# Live births (quadratic)" = "livebirths2", "Current Age" = "age"),
                   scale = TRUE) + apatheme + ggtitle("Levine Method")

#Plot for KDM
KDMplot1<-plot_summs(a3, b3,
                   coefs = c("# Live Births (linear)" = "livebirths", "# Live births (quadratic)" = "livebirths2", "Current Age" = "age"),
                   scale = TRUE) + apatheme + ggtitle("Klemera-Doubal Method")

#Plot for telo
TELOplot1<-plot_summs(a4, b4, 
                    coefs = c("# Live Births (linear)" = "livebirths", "# Live births (quadratic)" = "livebirths2", "Current Age" = "age"),
                    scale = TRUE) + apatheme + ggtitle("Telomere Length (log)")

HDplot1
LMplot1
KDMplot1
TELOplot1
```

##Specify survey designs: ever pregnant
```{r message=FALSE}
base <- svydesign(id      = ~SDMVPSU,
                          strata  = ~SDMVSTRA,
                          weights = ~WTMEC4YR,
                          nest    = TRUE,
                          data    = df)

nhanesDesign1<- subset(base, age > 17 & age <=84 & pregnant != 1 & everpreg > -1 & menopause == 0)
nhanesDesign2<- subset(base, age > 17 & age <=84 & pregnant != 1 & everpreg > -1 & menopause == 1)
```

##Most important model (all covariates, no interactions): Premenopausal, ever pregnant
```{r message=FALSE, echo=T, results='hide'}
summary(c1 <- svyglm(HD ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
summary(c2 <- svyglm(LM ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
summary(c3 <- svyglm(KDM ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))      
summary(c4 <- svyglm(log_telo ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign1))
```

##Tables for premenopausal women, ever pregnant
```{r, results = 'asis', echo = FALSE}
stargazer(c1,c2,c3,c4, single.row = F, align = T, intercept.bottom = T, no.space = T, dep.var.caption = "", type = "html", out = "../Output/Tables/models_ever_pre-meno.html", report = "vc*",   star.char = c("+", "*", "**", "***"), digits = 3, notes = "+p<0.1;*p<0.05;**p<0.01;***p<0.001 ", notes.append = FALSE)
```

##Most important model (all covariates, no interactions): Postmenopausal, ever pregnant
```{r message=FALSE, echo=T, results='hide'}
summary(d1 <- svyglm(HD ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
summary(d2 <- svyglm(LM ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
summary(d3 <- svyglm(KDM ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))      
summary(d4 <- svyglm(log_telo ~ everpreg + age + bmi + smoking + education + poverty + ridreth1 + food_security, design = nhanesDesign2))
```

##Tables for postmenopausal women, ever pregnant
```{r, results = 'asis', echo = FALSE}
stargazer(d1,d2,d3,d4, single.row = F, align = T, intercept.bottom = T, no.space = T, dep.var.caption = "", type = "html", out = "../Output/Tables/models_ever_post-meno.html", report = "vc*",   star.char = c("+", "*", "**", "***"), digits = 3, notes = "+p<0.1;*p<0.05;**p<0.01;***p<0.001 ", notes.append = FALSE)
```


```{r}
stargazer(c1,c2,c3,d1,d2,d3, 
          single.row = F, 
          align = T, 
          intercept.bottom = T, 
          no.space = T, 
          dep.var.caption = "Biological Aging (BA) Measures",
          dep.var.labels = c("", "Pre-menopausal", "", "", "Post-menopausal", ""),
          column.labels = c("HD", "LM", "KDM", "HD", "LM", "KDM"),
          covariate.labels = c("Births", "Births2", "Age",
                               "BMI", "Smoking (Y/N)", "Education",
                               "Poverty", "RIDRETH", "Food Security", "Intercept"),
          type = "html", 
          out = "../Output/Tables/models_ever_all-meno.html", 
          report = "vc*",   
          star.char = c("+", "*", "**", "***"), 
          digits = 3, 
          notes = "+p<0.1;*p<0.05;**p<0.01;***p<0.001 ", 
          notes.append = FALSE)

```








##Plots, ever pregnant (model 1 = premenopausal, model 2 = postmenopausal)
```{r message=FALSE, echo=FALSE}
#Plot for HD
HDplot2<-plot_summs(c1, d1, 
                   coefs = c("Parity" = "everpreg", "Current Age" = "age"), 
                   scale = TRUE) + apatheme + ggtitle("Homeostatic Dysregulation")

#Plot for LM
LMplot2<-plot_summs(c2, d2,
                   coefs = c("Parity" = "everpreg", "Current Age" = "age"),
                   scale = TRUE) + apatheme + ggtitle("Levine Method")

#Plot for KDM
KDMplot2<-plot_summs(c3, d3,  
                    coefs = c("Parity" = "everpreg", "Current Age" = "age"),
                    scale = TRUE) + apatheme + ggtitle("Klemera-Doubal Method")

#Plot for telo
TELOplot2<-plot_summs(c4, d4, 
                     coefs = c("Parity" = "everpreg", "Current Age" = "age"), 
                     scale = TRUE) + apatheme + ggtitle("Telomere Length (log)")

HDplot2
LMplot2
KDMplot2
TELOplot2
```

#Base models only adjusted for Chronological Age (stratified by menopausal status) 

```{r}
base <- svydesign(id      = ~SDMVPSU,
                          strata  = ~SDMVSTRA,
                          weights = ~WTMEC4YR,
                          nest    = TRUE,
                          data    = df)

nhanesDesign1<- subset(base, age > 17 & age <=84 & pregnant != 1 & livebirths > -1 & menopause == 0)
nhanesDesign2<- subset(base, age > 17 & age <=84 & pregnant != 1 & livebirths > -1 & menopause == 1)

#Premenopausal
summary(e1 <- svyglm(HD ~ livebirths + livebirths2 + age, design = nhanesDesign1))
summary(e2 <- svyglm(LM ~ livebirths + livebirths2 + age, design = nhanesDesign1))
summary(e3 <- svyglm(KDM ~ livebirths + livebirths2 + age, design = nhanesDesign1))      
summary(e4 <- svyglm(log_telo ~ livebirths + livebirths2 + age, design = nhanesDesign1))

#Postmenopausal
summary(f1 <- svyglm(HD ~ livebirths + livebirths2 + age, design = nhanesDesign2))
summary(f2 <- svyglm(LM ~ livebirths + livebirths2 + age, design = nhanesDesign2))
summary(f3 <- svyglm(KDM ~ livebirths + livebirths2 + age, design = nhanesDesign2))      
summary(f4 <- svyglm(log_telo ~ livebirths + livebirths2 + age, design = nhanesDesign2))
```

#Plots of Biological Age by Chronological Age across different numbers of pregnancies (Premenopausal)
Here are some different plots showing the relationship between chronological age and biological age for various numbers of births. For now we'll focus on graphs for mdoels only adjusting for chronological age. 

Extracting coefficients for each of the models. 

```{r}
coef_HD_pre <- e1$coefficients
coef_LM_pre <- e2$coefficients
coef_KDM_pre <- e3$coefficients
coef_Telo_pre <- e4$coefficients
```

Creating subsets holding individuals that would have all data in model 2 for each aging measure. 

```{r}
df_HD_pre <- subset(df, age>17 & age < 51 & pregnant !=1 & livebirths>-1 & HD >-5 & menopause == 0)
df_LM_pre <- subset(df, age>17 & age < 51 & pregnant !=1 & livebirths>-1 & LM >-11 & menopause == 0)
df_KDM_pre <- subset(df, age>17 & age < 51 & pregnant !=1 & livebirths>-1 & KDM >-5 & menopause == 0)
df_Telo_pre <- subset(df, age>17 & age < 51 & pregnant !=1 & livebirths>-1 & log_telo >-5 & menopause == 0)
```

Homeostatic dysregulation at 0,1,3,5,and 7 births. 

```{r}
fun_HD_0 <- function(x) {
  livebirths=0
  livebirths2=0
  coef_HD_pre[1] + coef_HD_pre[2]*livebirths + coef_HD_pre[3]*livebirths2 + coef_HD_pre[4]*x
}
fun_HD_1 <- function(x) {
  livebirths=1
  livebirths2=1
   coef_HD_pre[1] + coef_HD_pre[2]*livebirths + coef_HD_pre[3]*livebirths2 + coef_HD_pre[4]*x
}
fun_HD_3 <- function(x) {
  livebirths=3
  livebirths2=9
   coef_HD_pre[1] + coef_HD_pre[2]*livebirths + coef_HD_pre[3]*livebirths2 + coef_HD_pre[4]*x
}
fun_HD_5 <- function(x) {
  livebirths=5
  livebirths2=25
   coef_HD_pre[1] + coef_HD_pre[2]*livebirths + coef_HD_pre[3]*livebirths2 + coef_HD_pre[4]*x
}
fun_HD_7 <- function(x) {
  livebirths=7
  livebirths2=49
   coef_HD_pre[1] + coef_HD_pre[2]*livebirths + coef_HD_pre[3]*livebirths2 + coef_HD_pre[4]*x
}

ggplot(data=df_HD_pre, aes(x = age, y = HD)) + 
  ggtitle("Homeostatic Dysregulation and Chronological Age by Pregnancy Number (Pre)") +
  xlab("Chronological Age") + 
  ylab("Homeostatic Dysregulation")+
  stat_function(fun=fun_HD_0, color="green", size=2)+
  stat_function(fun=fun_HD_1, color="purple", size=2) +
  stat_function(fun=fun_HD_3, color="blue", size=2) +
  stat_function(fun=fun_HD_5, color="orange", size=2) +
  stat_function(fun=fun_HD_7, color="red", size=2) 
```
 
LM Biological Age at 1, 2, 3, and 4 births

```{r}
fun_LM_0 <- function(x) {
  livebirths=0
  livebirths2=0
  coef_LM_pre[1] + coef_LM_pre[2]*livebirths + coef_LM_pre[3]*livebirths2 + coef_LM_pre[4]*x
}
fun_LM_1 <- function(x) {
  livebirths=1
  livebirths2=1
   coef_LM_pre[1] + coef_LM_pre[2]*livebirths + coef_LM_pre[3]*livebirths2 + coef_LM_pre[4]*x
}
fun_LM_3 <- function(x) {
  livebirths=3
  livebirths2=9
   coef_LM_pre[1] + coef_LM_pre[2]*livebirths + coef_LM_pre[3]*livebirths2 + coef_LM_pre[4]*x
}
fun_LM_5 <- function(x) {
  livebirths=5
  livebirths2=25
   coef_LM_pre[1] + coef_LM_pre[2]*livebirths + coef_LM_pre[3]*livebirths2 + coef_LM_pre[4]*x
}
fun_LM_7 <- function(x) {
  livebirths=7
  livebirths2=49
   coef_LM_pre[1] + coef_LM_pre[2]*livebirths + coef_LM_pre[3]*livebirths2 + coef_LM_pre[4]*x
}

ggplot(data=df_LM_pre, aes(x = age, y = LM)) + 
  ggtitle("LM Biological Age and Chronological Age by Pregnancy Number (Pre") +
  xlab("Chronological Age") + 
  ylab("LM Biological Age")+
  stat_function(fun=fun_LM_0, color="green", size=2)+
  stat_function(fun=fun_LM_1, color="purple", size=2) +
  stat_function(fun=fun_LM_3, color="blue", size=2) +
  stat_function(fun=fun_LM_5, color="orange", size=2) +
  stat_function(fun=fun_LM_7, color="red", size=2) 
```

KDM Biological Age at 1, 2, 3, and 4 births

```{r}
fun_KDM_0 <- function(x) {
  livebirths=0
  livebirths2=0
  coef_KDM_pre[1] + coef_KDM_pre[2]*livebirths + coef_KDM_pre[3]*livebirths2 + coef_KDM_pre[4]*x
}
fun_KDM_1 <- function(x) {
  livebirths=1
  livebirths2=1
   coef_KDM_pre[1] + coef_KDM_pre[2]*livebirths + coef_KDM_pre[3]*livebirths2 + coef_KDM_pre[4]*x
}
fun_KDM_3 <- function(x) {
  livebirths=3
  livebirths2=9
   coef_KDM_pre[1] + coef_KDM_pre[2]*livebirths + coef_KDM_pre[3]*livebirths2 + coef_KDM_pre[4]*x
}
fun_KDM_5 <- function(x) {
  livebirths=5
  livebirths2=25
   coef_KDM_pre[1] + coef_KDM_pre[2]*livebirths + coef_KDM_pre[3]*livebirths2 + coef_KDM_pre[4]*x
}
fun_KDM_7 <- function(x) {
  livebirths=7
  livebirths2=49
   coef_KDM_pre[1] + coef_KDM_pre[2]*livebirths + coef_KDM_pre[3]*livebirths2 + coef_KDM_pre[4]*x
}

ggplot(data=df_KDM_pre, aes(x = age, y = KDM)) + 
  ggtitle("KDM Biological Age and Chronological Age by Pregnancy Number (Pre)") +
  xlab("Chronological Age") + 
  ylab("KDM Biological Age")+
  stat_function(fun=fun_KDM_0, color="green", size=2)+
  stat_function(fun=fun_KDM_1, color="purple", size=2) +
  stat_function(fun=fun_KDM_3, color="blue", size=2) +
  stat_function(fun=fun_KDM_5, color="orange", size=2) +
  stat_function(fun=fun_KDM_7, color="red", size=2) 
```

Telomere Length at 1, 2, 3, and 4 births

```{r}
fun_Telo_0 <- function(x) {
  livebirths=0
  livebirths2=0
  coef_Telo_pre[1] + coef_Telo_pre[2]*livebirths + coef_Telo_pre[3]*livebirths2 + coef_Telo_pre[4]*x
}
fun_Telo_1 <- function(x) {
  livebirths=1
  livebirths2=1
   coef_Telo_pre[1] + coef_Telo_pre[2]*livebirths + coef_Telo_pre[3]*livebirths2 + coef_Telo_pre[4]*x
}
fun_Telo_3 <- function(x) {
  livebirths=3
  livebirths2=9
   coef_Telo_pre[1] + coef_Telo_pre[2]*livebirths + coef_Telo_pre[3]*livebirths2 + coef_Telo_pre[4]*x
}
fun_Telo_5 <- function(x) {
  livebirths=5
  livebirths2=25
   coef_Telo_pre[1] + coef_Telo_pre[2]*livebirths + coef_Telo_pre[3]*livebirths2 + coef_Telo_pre[4]*x
}
fun_Telo_7 <- function(x) {
  livebirths=7
  livebirths2=49
   coef_Telo_pre[1] + coef_Telo_pre[2]*livebirths + coef_Telo_pre[3]*livebirths2 + coef_Telo_pre[4]*x
}

ggplot(data=df_Telo_pre, aes(x = age, y = log_telo)) + 
  ggtitle("Telomere Length and Chronological Age by Pregnancy Number (Pre)") +
  xlab("Chronological Age") + 
  ylab("Log Telomere Length")+
  stat_function(fun=fun_Telo_0, color="green", size=2)+
  stat_function(fun=fun_Telo_1, color="purple", size=2) +
  stat_function(fun=fun_Telo_3, color="blue", size=2) +
  stat_function(fun=fun_Telo_5, color="orange", size=2) +
  stat_function(fun=fun_Telo_7, color="red", size=2) 
```

#Plots of Biological Age by Chronological Age across different numbers of pregnancies (Postmenopausal)
Here are some different plots showing the relationship between chronological age and biological age for various numbers of births. For now we'll focus on graphs for mdoels only adjusting for chronological age. 

Extracting coefficients for each of the models. 

```{r}
coef_HD_post <- f1$coefficients
coef_LM_post <- f2$coefficients
coef_KDM_post <- f3$coefficients
coef_Telo_post <- f4$coefficients
```

Creating subsets holding individuals that would have all data in model 2 for each aging measure. 

```{r}
df_HD_post <- subset(df, age>17 & age < 51 & pregnant !=1 & livebirths>-1 & HD >-5 & menopause == 1)
df_LM_post <- subset(df, age>17 & age < 51 & pregnant !=1 & livebirths>-1 & LM >-11 & menopause == 1)
df_KDM_post <- subset(df, age>17 & age < 51 & pregnant !=1 & livebirths>-1 & KDM >-5 & menopause == 1)
df_Telo_post <- subset(df, age>17 & age < 51 & pregnant !=1 & livebirths>-1 & log_telo >-5 & menopause == 1)
```

Homeostatic dysregulation at 0,1,3,5,and 7 births. 

```{r}
fun_HD_0 <- function(x) {
  livebirths=0
  livebirths2=0
  coef_HD_post[1] + coef_HD_post[2]*livebirths + coef_HD_post[3]*livebirths2 + coef_HD_post[4]*x
}
fun_HD_1 <- function(x) {
  livebirths=1
  livebirths2=1
   coef_HD_post[1] + coef_HD_post[2]*livebirths + coef_HD_post[3]*livebirths2 + coef_HD_post[4]*x
}
fun_HD_3 <- function(x) {
  livebirths=3
  livebirths2=9
   coef_HD_post[1] + coef_HD_post[2]*livebirths + coef_HD_post[3]*livebirths2 + coef_HD_post[4]*x
}
fun_HD_5 <- function(x) {
  livebirths=5
  livebirths2=25
   coef_HD_post[1] + coef_HD_post[2]*livebirths + coef_HD_post[3]*livebirths2 + coef_HD_post[4]*x
}
fun_HD_7 <- function(x) {
  livebirths=7
  livebirths2=49
   coef_HD_post[1] + coef_HD_post[2]*livebirths + coef_HD_post[3]*livebirths2 + coef_HD_post[4]*x
}

ggplot(data=df_HD_post, aes(x = age, y = HD)) + 
  ggtitle("Homeostatic Dysregulation and Chronological Age by Pregnancy Number (Post)") +
  xlab("Chronological Age") + 
  ylab("Homeostatic Dysregulation")+
  stat_function(fun=fun_HD_0, color="green", size=2)+
  stat_function(fun=fun_HD_1, color="purple", size=2) +
  stat_function(fun=fun_HD_3, color="blue", size=2) +
  stat_function(fun=fun_HD_5, color="orange", size=2) +
  stat_function(fun=fun_HD_7, color="red", size=2) 
```
 
LM Biological Age at 1, 2, 3, and 4 births

```{r}
fun_LM_0 <- function(x) {
  livebirths=0
  livebirths2=0
  coef_LM_post[1] + coef_LM_post[2]*livebirths + coef_LM_post[3]*livebirths2 + coef_LM_post[4]*x
}
fun_LM_1 <- function(x) {
  livebirths=1
  livebirths2=1
   coef_LM_post[1] + coef_LM_post[2]*livebirths + coef_LM_post[3]*livebirths2 + coef_LM_post[4]*x
}
fun_LM_3 <- function(x) {
  livebirths=3
  livebirths2=9
   coef_LM_post[1] + coef_LM_post[2]*livebirths + coef_LM_post[3]*livebirths2 + coef_LM_post[4]*x
}
fun_LM_5 <- function(x) {
  livebirths=5
  livebirths2=25
   coef_LM_post[1] + coef_LM_post[2]*livebirths + coef_LM_post[3]*livebirths2 + coef_LM_post[4]*x
}
fun_LM_7 <- function(x) {
  livebirths=7
  livebirths2=49
   coef_LM_post[1] + coef_LM_post[2]*livebirths + coef_LM_post[3]*livebirths2 + coef_LM_post[4]*x
}

ggplot(data=df_LM_post, aes(x = age, y = LM)) + 
  ggtitle("LM Biological Age and Chronological Age by Pregnancy Number (Post") +
  xlab("Chronological Age") + 
  ylab("LM Biological Age")+
  stat_function(fun=fun_LM_0, color="green", size=2)+
  stat_function(fun=fun_LM_1, color="purple", size=2) +
  stat_function(fun=fun_LM_3, color="blue", size=2) +
  stat_function(fun=fun_LM_5, color="orange", size=2) +
  stat_function(fun=fun_LM_7, color="red", size=2) 
```

KDM Biological Age at 1, 2, 3, and 4 births

```{r}
fun_KDM_0 <- function(x) {
  livebirths=0
  livebirths2=0
  coef_KDM_post[1] + coef_KDM_post[2]*livebirths + coef_KDM_post[3]*livebirths2 + coef_KDM_post[4]*x
}
fun_KDM_1 <- function(x) {
  livebirths=1
  livebirths2=1
   coef_KDM_post[1] + coef_KDM_post[2]*livebirths + coef_KDM_post[3]*livebirths2 + coef_KDM_post[4]*x
}
fun_KDM_3 <- function(x) {
  livebirths=3
  livebirths2=9
   coef_KDM_post[1] + coef_KDM_post[2]*livebirths + coef_KDM_post[3]*livebirths2 + coef_KDM_post[4]*x
}
fun_KDM_5 <- function(x) {
  livebirths=5
  livebirths2=25
   coef_KDM_post[1] + coef_KDM_post[2]*livebirths + coef_KDM_post[3]*livebirths2 + coef_KDM_post[4]*x
}
fun_KDM_7 <- function(x) {
  livebirths=7
  livebirths2=49
   coef_KDM_post[1] + coef_KDM_post[2]*livebirths + coef_KDM_post[3]*livebirths2 + coef_KDM_post[4]*x
}

ggplot(data=df_KDM_post, aes(x = age, y = KDM)) + 
  ggtitle("KDM Biological Age and Chronological Age by Pregnancy Number (Post)") +
  xlab("Chronological Age") + 
  ylab("KDM Biological Age")+
  stat_function(fun=fun_KDM_0, color="green", size=2)+
  stat_function(fun=fun_KDM_1, color="purple", size=2) +
  stat_function(fun=fun_KDM_3, color="blue", size=2) +
  stat_function(fun=fun_KDM_5, color="orange", size=2) +
  stat_function(fun=fun_KDM_7, color="red", size=2) 
```

Telomere Length at 1, 2, 3, and 4 births

```{r}
fun_Telo_0 <- function(x) {
  livebirths=0
  livebirths2=0
  coef_Telo_post[1] + coef_Telo_post[2]*livebirths + coef_Telo_post[3]*livebirths2 + coef_Telo_post[4]*x
}
fun_Telo_1 <- function(x) {
  livebirths=1
  livebirths2=1
   coef_Telo_post[1] + coef_Telo_post[2]*livebirths + coef_Telo_post[3]*livebirths2 + coef_Telo_post[4]*x
}
fun_Telo_3 <- function(x) {
  livebirths=3
  livebirths2=9
   coef_Telo_post[1] + coef_Telo_post[2]*livebirths + coef_Telo_post[3]*livebirths2 + coef_Telo_post[4]*x
}
fun_Telo_5 <- function(x) {
  livebirths=5
  livebirths2=25
   coef_Telo_post[1] + coef_Telo_post[2]*livebirths + coef_Telo_post[3]*livebirths2 + coef_Telo_post[4]*x
}
fun_Telo_7 <- function(x) {
  livebirths=7
  livebirths2=49
   coef_Telo_post[1] + coef_Telo_post[2]*livebirths + coef_Telo_post[3]*livebirths2 + coef_Telo_post[4]*x
}

ggplot(data=df_Telo_post, aes(x = age, y = log_telo)) + 
  ggtitle("Telomere Length and Chronological Age by Pregnancy Number (Post)") +
  xlab("Chronological Age") + 
  ylab("Log Telomere Length")+
  stat_function(fun=fun_Telo_0, color="green", size=2)+
  stat_function(fun=fun_Telo_1, color="purple", size=2) +
  stat_function(fun=fun_Telo_3, color="blue", size=2) +
  stat_function(fun=fun_Telo_5, color="orange", size=2) +
  stat_function(fun=fun_Telo_7, color="red", size=2) 
```
