---
title: "Residuals"
author: "Waylon J. Hastings"
date: "10/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Loading Libraries and datafiles 

```{r}
library(haven)
library(psych)
library(dplyr)
library(ggplot2)
library(survey)
library(foreign)
library(psych)
library(jtools)
library(car)

# NHANES <- read.csv("~/Documents/GitHub/NHANES_CoR/Data/NHANES_090920.csv", header=TRUE)
NHANES <- read.csv("~/GitHub/NHANES_CoR/Data/NHANES_090920.csv", header=TRUE)
# load("~/Documents/GitHub/NHANES_CoR/Data/NHANES_BA_Talia.rda")
load("~/GitHub/NHANES_CoR/Data/NHANES_BA_Talia.rda")
```

#Housekeeping 

```{r}
df <- merge (NHANES, BA_NHANES_Parity, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="PhenoAge"] <- "LM"
df$livebirths2 <- df$livebirths^2
df$BMI2 <- df$BMI^2
df$monthssincebirth <- as.numeric(df$monthssincebirth)

#Make categorical BMI variable
df$BMI_cat <- car::recode(df$BMI, "
                          0:18.499 = 'underweight';
                          18.50:24.99 = 'normal';
                          25.0:29.99 = 'overweight';
                          30.0:500 = 'obesity'")
#Make race/ethnicity factor variable
df$RIDRETH1 <- as.factor(df$RIDRETH1)
df <- within(df, RIDRETH1 <- relevel(RIDRETH1, ref = "0"))
#Make smoking factor variable
df$smoking <- as.factor(df$smoking)
df <- within(df, smoking <- relevel(smoking, ref = "0"))
#Make education factor variable
df$DMDEDUC2 <- as.factor(df$DMDEDUC2)
df <- within(df, DMDEDUC2 <- relevel(DMDEDUC2, ref = "5"))
#Make menopause factor
df$menopause <- as.factor(df$menopause)
df <- within(df, menopause <- relevel(menopause, ref = "0"))
#Make livebirths_dichot factor
df$livebirths_dichot <- as.factor(df$livebirths_dichot)
df <- within(df, livebirths_dichot <- relevel(livebirths_dichot, ref = "0"))
#Update RIDEXPRG variable
df <- df %>%
  mutate(RIDEXPRG = case_when(RIDEXPRG == 3 ~3, RIAGENDR == 2 & RIDAGEYR > 59 & RIDAGEYR <= 84 ~ 2, RIDEXPRG == 1 ~ 1, RIDEXPRG == 2 ~ 2)) %>%
  mutate_at(vars(RIDEXPRG), funs(as.factor))
```

#First Full Residuals: Biosample (N=5,870)

```{r}
#Make LM resid variable
df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
lm2 <- lm(LM~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,52)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,52)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,52)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,52)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid"
```

#First Split Residuals: Biosample by menopause status 

```{r} 
#Make LM resid variable
df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1 & menopause==0)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
lm2 <- lm(LM~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,56)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid_premeno"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,56)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid_premeno"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,56)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid_premeno"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,56)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid_premeno"

df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1 & menopause==1)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))

#Make LM resid variable
lm2 <- lm(LM~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,60)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid_postmeno"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,60)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid_postmeno"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,60)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid_postmeno"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df3)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df3, resids)
resids2 <- resids[,c(1,60)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid_postmeno"
```

#Second Full Residuals: Primary Sample (N=4,418)

```{r}
df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
df4 <- filter(df3, livebirths < 8)
df5 <- df4 %>% filter(complete.cases(RIDAGEYR, BMI, INDFMPIR, smoking, DMDEDUC2, RIDRETH1, livebirths, menopause, LM))

#Make LM resid variable
lm2 <- lm(LM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,64)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid2"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,64)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid2"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,64)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid2"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,64)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid2"
```

#Second Split Residuals: Primary Sample by Menopause Status 

```{r}
df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1 & menopause==0)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
df4 <- filter(df3, livebirths < 8)
df5 <- df4 %>% filter(complete.cases(RIDAGEYR, BMI, INDFMPIR, smoking, DMDEDUC2, RIDRETH1, livebirths, menopause, LM))

#Make LM resid variable
lm2 <- lm(LM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,68)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid2_premeno"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,68)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid2_premeno"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,68)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid2_premeno"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,68)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid2_premeno"

df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1 & menopause==1)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
df4 <- filter(df3, livebirths < 8)
df5 <- df4 %>% filter(complete.cases(RIDAGEYR, BMI, INDFMPIR, smoking, DMDEDUC2, RIDRETH1, livebirths, menopause, LM))

#Make LM resid variable
lm2 <- lm(LM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,72)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid2_postmeno"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,72)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid2_postmeno"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,72)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid2_postmeno"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,72)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid2_postmeno"
```

#Third Full Residuals: Sensitivity Analysis Sample 1 (N=5,184)

```{r}
df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
df4 <- filter(df3, livebirths < 8)
lm2 <- lm(LM~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,76)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid3"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,76)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid3"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,76)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid3"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,76)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid3"
```

#Third Split Residuals: Sensitivity Analysis Sample 1 by Menopause Status 

```{r} 
df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1 & menopause==0)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
df4 <- filter(df3, livebirths < 8)
lm2 <- lm(LM~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,80)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid3_premeno"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,80)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid3_premeno"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,80)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid3_premeno"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,80)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid3_premeno"

df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1 & menopause==1)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
df4 <- filter(df3, livebirths < 8)

#Make LM resid variable
lm2 <- lm(LM~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,84)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid3_postmeno"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,84)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid3_postmeno"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,84)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid3_postmeno"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df4)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df4, resids)
resids2 <- resids[,c(1,84)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid3_postmeno"
```

#Fourth Full Residuals: Sensitivity Analysis 2 Sample (N=3,587)

```{r}
df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
df4 <- filter(df3, livebirths < 8)
df5 <- df4 %>% filter(complete.cases(RIDAGEYR, BMI, INDFMPIR, smoking, DMDEDUC2, RIDRETH1, livebirths, menopause, LM,yearssincelastbirth))

#Make LM resid variable
lm2 <- lm(LM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,88)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid4"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,88)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid4"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,88)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid4"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,88)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid4"
```

#Fourth Split Residuals: Sensitivity Analysis 2 Sample by Menopause Status (N=3,587) 

```{r}
df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1 & menopause==0)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
df4 <- filter(df3, livebirths < 8)
df5 <- df4 %>% filter(complete.cases(RIDAGEYR, BMI, INDFMPIR, smoking, DMDEDUC2, RIDRETH1, livebirths, menopause, LM,yearssincelastbirth))

#Make LM resid variable
lm2 <- lm(LM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,92)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid4_premeno"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,92)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid4_premeno"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,92)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid4_premeno"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,92)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid4_premeno"

df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1 & menopause==1)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
df4 <- filter(df3, livebirths < 8)
df5 <- df4 %>% filter(complete.cases(RIDAGEYR, BMI, INDFMPIR, smoking, DMDEDUC2, RIDRETH1, livebirths, menopause, LM,yearssincelastbirth))

#Make LM resid variable
lm2 <- lm(LM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,96)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "LM_age_resid4_postmeno"

#Make KDM resid variable
lm2 <- lm(KDM~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,96)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "KDM_age_resid4_postmeno"

#Make HD resid variable
lm2 <- lm(LOG_HD~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,96)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "HD_age_resid4_postmeno"

#Make AL resid variable
lm2 <- lm(AL~RIDAGEYR, data = df5)
resids <- as.data.frame(lm2$residuals)
resids <- cbind(df5, resids)
resids2 <- resids[,c(1,96)]
df <- merge (df, resids2, by = "SEQN", all = TRUE)
colnames(df)[colnames(df)=="lm2$residuals"] <- "AL_age_resid4_postmeno"
```

#Survey Specs

```{r}
#Create correct sample weights; see https://wwwn.cdc.gov/nchs/nhanes/tutorials/module3.aspx
df <- df %>%
  mutate(WTMEC12YR = NA,
         WTMEC12YR = ifelse(SDDSRVYR == 1 | SDDSRVYR == 2, (WTMEC4YR/3), WTMEC12YR),
         WTMEC12YR = ifelse(SDDSRVYR == 3 | SDDSRVYR == 4 | SDDSRVYR == 5 | SDDSRVYR == 6, (WTMEC2YR/6), WTMEC12YR))

#write this df to .dta
#write.dta(df, "NHANES_repro_file_101620.dta")

#Set survey specs
base <- svydesign(id      = ~SDMVPSU,
                  strata  = ~SDMVSTRA,
                  weights = ~WTMEC12YR,
                  nest    = TRUE,                  
                  data    = df)

#Excluding 8+ live births
nhanesDesign1<- subset(base, RIDAGEYR > 17 & RIDAGEYR <=84 & RIAGENDR == 2 & livebirths > -1 & livebirths < 8 & RIDEXPRG == 2)


```


#Full Sample Analyses 


```{r}
#Primary Analyses: All covariates 
summ(LM1 <- svyglm(LM_age_resid2 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid2 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid2 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid2 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)

#Sensitivity Analyses 1a: Age covariate only
summ(LM1 <- svyglm(LM_age_resid3 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid3 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid3 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid3 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)

#Sensitivity Analyses 1b: Age covariate only, but with primary analytical sample 
nhanesDesign1<- subset(base, RIDAGEYR > 17 & RIDAGEYR <=84 & RIAGENDR == 2 & livebirths > -1 & livebirths < 8 & RIDEXPRG == 2 & BMI > 0 & INDFMPIR > -1 & !(is.na(smoking)) & !(is.na(DMDEDUC2)))

summ(LM1 <- svyglm(LM_age_resid2 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid2 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid2 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid2 ~ livebirths*menopause + livebirths2*menopause + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)

#Sensitivity Analyses 2: Years since last birth 
#Excluding 8+ live births
nhanesDesign1<- subset(base, RIDAGEYR > 17 & RIDAGEYR <=84 & RIAGENDR == 2 & livebirths > -1 & livebirths < 8 & RIDEXPRG == 2)

summ(LM1 <- svyglm(LM_age_resid4 ~ livebirths*menopause + livebirths2*menopause + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid4 ~ livebirths*menopause + livebirths2*menopause + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid4 ~ livebirths*menopause + livebirths2*menopause + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid4 ~ livebirths*menopause + livebirths2*menopause + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 4)
```


#Stratified Analyses: Premenopausal Women


```{r}
#Primary analysis: All Covariates 
nhanesDesign1<- subset(base, RIDAGEYR > 17 & RIDAGEYR <=84 & RIAGENDR == 2 & livebirths > -1 & livebirths < 8 & RIDEXPRG == 2 & menopause ==0)

summ(LM1 <- svyglm(LM_age_resid2_premeno ~ livebirths + livebirths2 + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid2_premeno ~ livebirths + livebirths2 + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid2_premeno ~ livebirths + livebirths2 + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid2_premeno ~ livebirths + livebirths2 + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)

#Sensitivity Analysis 1: Age covariate only
summ(LM1 <- svyglm(LM_age_resid3_premeno ~ livebirths + livebirths2 + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid3_premeno ~ livebirths + livebirths2 + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid3_premeno ~ livebirths + livebirths2 + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid3_premeno ~ livebirths + livebirths2 + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)

#Sensitivity Analysis 2: Years since last birth 
summ(LM1 <- svyglm(LM_age_resid4_premeno ~ livebirths + livebirths2 + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid4_premeno ~ livebirths + livebirths2 + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid4_premeno ~ livebirths + livebirths2 + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid4_premeno ~ livebirths + livebirths2 + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 4)

#Sensitivity analysis 3: Months since lasts birth 
summ(LM1 <- svyglm(LM_age_resid4_premeno ~ livebirths*monthssincebirth + livebirths2*monthssincebirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid4_premeno ~ livebirths*monthssincebirth + livebirths2*monthssincebirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid4_premeno ~ livebirths*monthssincebirth + livebirths2*monthssincebirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid_premeno ~ livebirths*monthssincebirth + livebirths2*monthssincebirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 4)
```


#Stratified Analyses: Postmenopausal Women 


```{r}
nhanesDesign1<- subset(base, RIDAGEYR > 17 & RIDAGEYR <=84 & RIAGENDR == 2 & livebirths > -1 & livebirths < 8 & RIDEXPRG == 2 & menopause==1)

#Primary analysis: All covariates 
summ(LM1 <- svyglm(LM_age_resid2_postmeno ~ livebirths + livebirths2 + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid2_postmeno ~ livebirths + livebirths2 + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid2_postmeno ~ livebirths + livebirths2 + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid2_postmeno ~ livebirths + livebirths2 + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)

#Sensitivity Analysis 1: Age covariate only
summ(LM1 <- svyglm(LM_age_resid3_postmeno ~ livebirths + livebirths2 + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid3_postmeno ~ livebirths + livebirths2 + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid3_postmeno ~ livebirths + livebirths2 + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid3_postmeno ~ livebirths + livebirths2 + RIDAGEYR, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)

#Sensitivity Analysis 2: Years since last birth 
summ(LM1 <- svyglm(LM_age_resid4_postmeno ~ livebirths + livebirths2 + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid4_postmeno ~ livebirths + livebirths2 + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid4_postmeno ~ livebirths + livebirths2 + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid4_postmeno ~ livebirths + livebirths2 + livebirths*yearssincelastbirth + livebirths2*yearssincelastbirth + RIDAGEYR + BMI + BMI2 + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 4)
```

#Demographics by Menopausal Status 

```{r}
df2 <- filter(df, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2)
df3 <- df2 %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, LM))
df4 <- filter(df3, livebirths < 8)
df5 <- df4 %>% filter(complete.cases(RIDAGEYR, BMI, INDFMPIR, smoking, DMDEDUC2, RIDRETH1, livebirths, menopause, LM))
df5 <- dplyr::select(df5,SEQN,menopause,RIDAGEYR,BMI,INDFMPIR,smoking,DMDEDUC2,RIDRETH1,livebirths,livebirths_dichot,LM_age_resid2,HD_age_resid2,KDM_age_resid2,AL_age_resid2,LM,LOG_HD,KDM,AL)

#Continuous variables
describeBy(df5,df5$menopause)
#Categorical variables 
table(df5$smoking,df5$menopause)
table(df5$DMDEDUC2,df5$menopause)
table(df5$RIDRETH1,df5$menopause)
table(df5$livebirths_dichot,df5$menopause)

#Chi-Square tests
chisq.test(table(df5$smoking,df5$menopause))
chisq.test(table(df5$DMDEDUC2,df5$menopause))
chisq.test(table(df5$RIDRETH1,df5$menopause))
chisq.test(table(df5$livebirths_dichot,df5$menopause))

#t-tests
t.test(df5$RIDAGEYR~df5$menopause)
t.test(df5$BMI~df5$menopause)
t.test(df5$INDFMPIR~df5$menopause)
t.test(df5$livebirths~df5$menopause)
t.test(df5$LM_age_resid2~df5$menopause)

#Regression for effect of menopause on biological age with adjustment for age, BMI, poverty, smoking, education, and ethnicity.

nhanesDesign1<- subset(base, RIDAGEYR > 17 & RIDAGEYR <=84 & RIAGENDR == 2 & livebirths > -1 & livebirths < 8 & RIDEXPRG == 2)

summ(LM1 <- svyglm(LM_age_resid2 ~ menopause + RIDAGEYR + BMI + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(HD1 <- svyglm(HD_age_resid2 ~ menopause + RIDAGEYR + BMI + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(KDM1 <- svyglm(KDM_age_resid2 ~ menopause + RIDAGEYR + BMI + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
summ(AL1 <- svyglm(AL_age_resid2 ~ menopause + RIDAGEYR + BMI + INDFMPIR + smoking + DMDEDUC2 + RIDRETH1, design = nhanesDesign1, data = df), confint = TRUE, digits = 3)
```

#Plots 

**Number of Live Births Histogram**

```{r}
df2 <- df %>% filter(complete.cases(livebirths, menopause, RIDAGEYR, BMI, INDFMPIR, RIDRETH1, smoking, DMDEDUC2,LM,KDM,LOG_HD))
df3 <- filter(df2, RIDAGEYR>17 & RIDAGEYR <= 84 & RIDEXPRG == 2 & RIAGENDR == 2 & livebirths >-1 & livebirths < 8)

df3$livebirths<-as.factor(df3$livebirths)

ggplot(df3, aes(x=livebirths, fill = menopause)) +
  geom_histogram(position = "dodge", stat = "count", color = "black") +
  theme_minimal() + scale_fill_grey() + 
  ylab("Frequency") + xlab("Live births") + 
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18)) +
  theme(legend.position = "none")
```


**Correlation Plots**

```{r}
df2 <- filter(df, RIAGENDR == 2 & RIDAGEYR > 17 & RIDAGEYR <= 84 & RIDEXPRG == 2)
df3 <- filter(df2, LM > 0)
df4 <- filter(df3, livebirths < 8)
df5 <- df4 %>% filter(complete.cases(RIDAGEYR, BMI, INDFMPIR, smoking, DMDEDUC2, RIDRETH1, livebirths, menopause, LM))

#Raw Measures
dfcor<-df5[c("RIDAGEYR", "LM", "LOG_HD", "KDM", "AL")]
dfcor <- dfcor %>% rename(
  "Age" = RIDAGEYR,
  "HD (log)" = LOG_HD
)

pairs.panels(dfcor, stars= TRUE, method = "pearson", hist.col = "gray", cex.labels=1.9, ellipses = FALSE) 

#Age-Adjusted Measures: Primary Sample 
dfcor<-df5[c("RIDAGEYR", "LM_age_resid2", "HD_age_resid2", "KDM_age_resid2", "AL_age_resid2")]
dfcor <- dfcor %>% rename(
  "Age" = RIDAGEYR,
  "HD (adjusted)" = HD_age_resid2,
  "LM (adjusted)" = LM_age_resid2,
  "KDM (adjusted)" = KDM_age_resid2,
  "AL (adjusted)" = AL_age_resid2
)

pairs.panels(dfcor, stars= TRUE, method = "pearson", hist.col = "gray", cex.labels=1.9, ellipses = FALSE)
```

**Marginsplot**

```{r}
margins <- read_excel("~/GitHub/NHANES_CoR/Data/marginsplot values 101620.xlsx")

pd <- position_dodge(0.15) 
margins$Group <- as.factor(margins$Group)

LM <- margins %>%
  filter(Measure == "LM") %>%
  ggplot(., aes(x =`Live births`, y = Point, group = Group, color = Group)) +
  geom_line(size = 1.4) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Point - SE, ymax = Point + SE), width = 0.1, position = pd, size = 1) +
  theme_classic() + 
  scale_color_grey() + 
  ylab("LM age acceleration")  +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18)) +
  scale_x_continuous(breaks = c(0:7), labels = c(0:7)) +
  theme(legend.position = "none")

KDM <- margins %>%
  filter(Measure == "KDM") %>%
  ggplot(., aes(x =`Live births`, y = Point, group = Group, color = Group)) +
  geom_line(size = 1.4) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Point - SE, ymax = Point + SE), width = 0.1, position = pd, size = 1) +
  theme_classic() + 
  scale_color_grey() + 
  ylab("KDM age acceleration")  +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18)) +
  scale_x_continuous(breaks = c(0:7), labels = c(0:7)) +
  theme(legend.position = "none")

HD <- margins %>%
  filter(Measure == "HD") %>%
  ggplot(., aes(x =`Live births`, y = Point, group = Group, color = Group)) +
  geom_line(size = 1.4) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Point - SE, ymax = Point + SE), width = 0.1, position = pd, size = 1) +
  theme_classic() + 
  scale_color_grey() + 
  ylab("HD age acceleration")  +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18)) +
  scale_x_continuous(breaks = c(0:7), labels = c(0:7)) +
  theme(legend.position = "none")

AL <- margins %>%
  filter(Measure == "AL") %>%
  ggplot(., aes(x =`Live births`, y = Point, group = Group, color = Group)) +
  geom_line(size = 1.4) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Point - SE, ymax = Point + SE), width = 0.1, position = pd, size = 1) +
  theme_classic() + 
  scale_color_grey() + 
  ylab("AL age acceleration")  +
  theme(axis.text = element_text(size = 14), axis.title = element_text(size = 18)) +
  scale_x_continuous(breaks = c(0:7), labels = c(0:7), limits = c(0,7)) +
  theme(legend.position = "none")


```